\textbf{Advection versus diffusion in intraventricular transport}

### 2025-03-25 paper/main.tex
fix fig1 references

We assume that viscous forces are dominating the intraventricular CSF flow,
and model the flow by the Stokes equations.
The net forces imparted by motile cilia on the CSF are modeled as a steady
bidirectional traction acting on parts of the ventricular wall
(\Crefpanels{fig:fig1}{d}), and we assume free velocity slip at the walls.
Simulations show that this traction alone
leads to a partial compartmentalization of the ventricular system,
with large-scale vortex structures in the anterior, middle and
posterior ventricles (\Crefpanels{fig:fig1}{e}). The highest flow
speeds ($26.7$ \textmu m/s) occur in the vicinity of the dorsal cilia
in the middle ventricle. To model the cardiac pulsations, we impose a
pulsatile pressure difference between the anterior and posterior
boundaries (\Crefpanels{fig:fig1}{d, \fixme{g}}). The resulting flow
is pulsatile and directional (\Crefpanels{fig:fig1}{\fixme{f}}), and
in the absence of the cilia, the vortex structures vanish. The
velocity magnitude peaks at $4.8$ \textmu m/s. CSF initially flows
rostrocaudally before changing direction to caudorostral in the middle
of the cardiac cycle.

>>>>

We assume that viscous forces are dominating the intraventricular CSF flow,
and model the flow by the Stokes equations.
The net forces imparted by motile cilia on the CSF are modeled as a steady
bidirectional traction acting on parts of the ventricular wall
(\Crefpanels{fig:fig1}{d}), and we assume free velocity slip at the walls.
Simulations show that this traction alone
leads to a partial compartmentalization of the ventricular system,
with large-scale vortex structures in the anterior, middle and
posterior ventricles (\Crefpanels{fig:fig1}{e}). The highest flow
speeds ($26.7$ \textmu m/s) occur in the vicinity of the dorsal cilia
in the middle ventricle. To model the cardiac pulsations, we impose a
pulsatile pressure difference between the anterior and posterior
boundaries (\Crefpanels{fig:fig1}{d, f}). The resulting flow
is pulsatile and directional (\Crefpanels{fig:fig1}{g}), and
in the absence of the cilia, the vortex structures vanish. The
velocity magnitude peaks at $4.8$ \textmu m/s. CSF initially flows
rostrocaudally before changing direction to caudorostral in the middle
of the cardiac cycle.
